---
title: "HW1"
author: "Yahan Xue"
date: "2026-09-23"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
I load 'penguins' dataset from the 'palmerpenguins' package.
```{r load_penguins}
data("penguins", package = "palmerpenguins")
```
The `penguins` dataset contains measurements of penguins, including variables
such as species, island, bill length, bill depth, flipper length, body mass,
sex, and year.

The dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

Next, I create a scatterplot of flipper length versus bill length, with points
colored according to penguin species.

```{r penguin_plot}
library(ggplot2)

penguin_plot =
  ggplot(
    penguins,
    aes(x = bill_length_mm, y = flipper_length_mm, color = species)
  ) +
  geom_point()

penguin_plot
```
```{r save_plot}
ggsave("penguin_scatterplot.png", plot = penguin_plot)
```


# Problem 2

I create a dataframe containing a random sample from a standard normal
distribution, a logical vector, a character vector, and a factor vector.

```{r create_dataframe}
library(tidyverse)

set.seed(123)

problem2_df =
  tibble::tibble(
    random_sample = rnorm(10),
    logical_vector = random_sample > 0,
    character_vector = c(
      "a", "b", "c", "d", "e",
      "f", "g", "h", "i", "j"
    ),
    factor_vector = factor(
      c(
        "level1", "level2", "level3", "level1", "level2",
        "level3", "level1", "level2", "level3", "level1"
      )
    )
  )
```

Next, I calculate the mean of each variable.

```{r calculate_means}
mean(pull(problem2_df, random_sample))
mean(pull(problem2_df, logical_vector))
mean(pull(problem2_df, character_vector))
mean(pull(problem2_df, factor_vector))
```
The mean can be calculated for the numeric and logical variables. For the
logical variable, R treats `TRUE` as 1 and `FALSE` as 0, so its mean represents
the proportion of values that are `TRUE`. The character and factor variables
do not have meaningful arithmetic means, and `mean()` returns `NA`
with a warning.

Next, I mannually convert the logical, character, and factor variables to
numeric variables.

```{r coercion, results='hide'}
as.numeric(pull(problem2_df, logical_vector))
as.numeric(pull(problem2_df, character_vector))
as.numeric(pull(problem2_df, factor_vector))
```

Logical values convert `FALSE` to 0 and `TRUE` to 1. So the mean of a logical variable can be calculated.

The character variable contains non-numeric strings, so it turns to numeric produces `NA` values.

The factor variable is converted to integer codes corresponding to its factor levels. These integer codes represent the internal coding of the factor levels, which are not meaningful numerical values. 